---
title: "Prepare data"
subtitle: "NYU Furman Center"
date: "`r Sys.Date()`"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r load-packages-set-options}
library(tidyverse) # general data manipulation and graphing
library(geoclient) # geocoding tool
library(DBI) # connecting R to database management systems

con <- furmanr::fcdb_connect()

geoclient_api_keys(
  Sys.getenv("GEOCLIENT_APP_ID"),
  Sys.getenv("GEOCLIENT_APP_KEY")
)

```

### Data Sources: Hotel Properties

We retrieve data from the [New York City Department of Finance](https://www1.nyc.gov/site/finance/index.page), which assesses each property in New York City for the purpose of, among other things, calculating property taxes. Using Property Tax System data, this we focus on properties with hotel building classifications as defined below. To account for hotels located on lots that also contain condos, we use a crosswalk file to obtain all BBLs associated with condo lots. We use the non-condo BBls to join in geographic information from MapPLUTO. 

* The hotel classes outside of H1-H8, including boutique hotels, hostels, extended stay hotels, and SROs (HR) were combined into the category we've titled "Miscellaneous". We exclude H5, H6, H7, and H8 - private club, apartment hotels, apartment hotel-coops, and dormitories - from our analysis.

* We retrieved union status information from the [Hotel Trades Council](https://hotelworkers.org/#find-union-hotels). Addresses were run through [the City of New York's geocoder](https://developer.cityofnewyork.us/api/geoclient-api) to obtain a mapable BBL. This union status is indicated in the data by `is_union`.

* Under Article 1 Chapter 5, certain hotels are eligible for conversion to residential use notwithstanding otherwise relevant regulations. This potential eligibility is indicated in the data by `article1_eligible`.
```{r raw-rpad-xwalk}
# reading in raw rpad data 
rpad_raw <- tbl(con, "rpad_raw_with_pts") %>%
  filter(
    str_detect(bldgcl,'^H.*') |  bldgcl == 'RH',
    year4 == '2020'
  ) 

xwalk <- tbl(con, "rpad_condo_xwalk") %>%
  filter(year =='2018') %>%
  transmute(
    bbl = bbl_unit,
    bbl_condo
  ) 

raw <- rpad_raw %>%
  filter(
    year4 == '2020'
  ) %>%
  select(
    bldgcl,
    bbl
  ) %>%
  left_join(xwalk, by = "bbl")%>%
  collect()
```

```{r raw-union}
# reading in raw union data, geocoding
unionhotel_raw <- read_csv("data/unionhotels.csv") %>%
  filter(state == "NY") %>%
  unite(hotel_addr, address, zip, sep = " ,",remove = FALSE)

unionhotel_geocoded <- unionhotel_raw %>%
  geo_search_data(hotel_addr) %>%
  filter(!no_results)%>%
  select(bbl, input_location)%>%
  rename(hotel_addr= input_location) %>%
  right_join(unionhotel_raw, by="hotel_addr")%>%
  select(bbl)%>%
  mutate(is_union = TRUE)%>%
  filter(!is.na(bbl))

article1_cds <-
  c(101, 102, 103, 104, 105, 106, 301, 302, 306, 307, 401, 402) # eligible community districts under art 1 chap 5
```

### Data Sources: Hotel Room Counts

The room count in our data, `final_rooms`, comes from two main sources: the Department of Finance's Notice of Property Value records and Hotels.com. The Notice of Property Value (NOPV) informs every property owner of the Department of Finance's assessment of property for the coming tax year. Also listed in these publicly available PDFs are hotel room counts for most hotel classified properties. These files are available through the [Department of Finance's Property Tax Public Access web portal](https://a836-pts-access.nyc.gov/care/forms/htmlframe.aspx?mode=content/home.htm).  

* Some January 2021 NOPVs did not contain room counts, though the previous year's NOPV did. In these cases we resorted to January 2020 NOPV room counts if available.

* For lots containing condos, we took the sum total number of hotel rooms located on that lot. This allows us to capture the total number of hotel rooms on each mappable lot, though there may actually be more than one distinct hotel property on that lot. All condo BBLs are listed in `condo_bbls`.

```{r load-nopv}
# reading in raw nopv data, combining
nopv21 <- read_csv("J:/DEPT/REUP/Projects/Hotels/nopv/dof-nopv-hotels-20210115.csv") %>%
  transmute(bbl = as.character(bbl),
            hotel_class,
            nopv21_rooms= ifelse(hotel_rooms < 1, NA_real_, hotel_rooms))

nopv20 <- read_csv("J:/DEPT/REUP/Projects/Hotels/nopv/dof-nopv-hotels-20200115.csv") %>%
  transmute(bbl = as.character(bbl),
            nopv20_rooms= ifelse(hotel_rooms < 1, NA_real_, hotel_rooms))

nopv_combined <- nopv20 %>%
  right_join(nopv21, by= "bbl")
```

 
```{r assign-nopv}
# assigning nopv room counts to lots
rpad_hotels <-  raw %>%
  left_join(nopv_combined, by = "bbl") %>% 
  mutate(nopv_rooms = coalesce(nopv21_rooms, nopv20_rooms))%>%
  select(-nopv21_rooms, -nopv20_rooms) %>%
  group_by(bbl_condo) %>%  
  mutate(
    map_nopv_rooms = case_when(
      is.na(bbl_condo) ~ nopv_rooms, # no condos on main lot
      TRUE ~ sum(unique(nopv_rooms), na.rm = TRUE) # condos on main lot 
    ),
    map_bbl = ifelse(is.na(bbl_condo), bbl, bbl_condo)
  ) %>% 
  ungroup() %>%
  distinct(map_bbl,nopv_rooms, .keep_all = TRUE) %>%
  group_by(map_bbl) %>% 
  mutate(
    condo_bbls = paste(bbl, collapse = ";"),
    num_condo_bbls = n_distinct(bbl)
  ) %>% 
  ungroup() %>% 
  distinct(map_bbl, .keep_all = TRUE) %>%
  select(-bbl, -bbl_condo, -nopv_rooms) %>% 
  rename(bbl = map_bbl, nopv_rooms = map_nopv_rooms)

dbWriteTable(con, "rpad_hotels", rpad_hotels, temporary=TRUE, overwrite = TRUE)

# combine MapPLUTO geographic data
rpad_with_pluto <- dbGetQuery(con, "
                              SELECT bldgclass, address, borough, cd, zonedist1, yearbuilt, builtfar, latitude, longitude, r.*
                              FROM mappluto_21v1 p
                              RIGHT JOIN rpad_hotels r
                                USING(bbl)
                              ")
```

While NOPV room counts are available for most hotel classified properties, an additional step of filling in missing room counts was made by scraping information from Hotels.com. Available in the dataset from this scrape are: `hotel_id` `hotel_rooms` `hotel_close_start` and `hotel_close_end`. Scraped hotel addresses were also run through [the City of New York's geocoder](https://developer.cityofnewyork.us/api/geoclient-api) to obtain a mappable BBL. Note that the method of geocoding addresses pulled from Hotels.com is imperfect, and there is a rough approximation of rooms at these addresses for which NOPV counts are missing. New hotels or hotels under construction in particular may not have accurate counts until completion. Ultimately, some properties still needed to be searched manually to obtain room counts, and in these cases source information is included in the `source` column. 

The final room count is determined first by the availability of January 2021 NOPV counts, and substituted from January 2020 NOPV counts and Hotels.com data when needed. 

```{r manual-search}
# reading in manual searches
missing_info <- read_csv("data/manual-searches.csv") %>%
  transmute(bbl = as.character(bbl),
            include,
            roomcount,
            source)
```

```{r}
# combining frames, hardcoding issues, filtering classes of interest
final_hotels <-
  dbGetQuery(
    con,
    "SELECT bbl, hotel_id, hotel_rooms, hotel_close_start, hotel_close_end FROM hotels_scraped_clean "
  ) %>%
  right_join(rpad_with_pluto, by = "bbl") %>%
  left_join(missing_info, by = "bbl") %>%
  left_join(unionhotel_geocoded, by = "bbl") %>% 
  mutate(
    yearbuilt = case_when(
      hotel_id == "ho1227757728" ~ 2018, # hardcoding hotel missing year built
      yearbuilt == 0 ~ NA_real_,
      TRUE ~ as.numeric(yearbuilt)
    ),
    include = ifelse(is.na(include), TRUE, include),
    is_union = ifelse(is.na(is_union), FALSE, is_union),
    nopv_rooms = ifelse(nopv_rooms == 0, NA_real_, nopv_rooms),
    final_rooms = coalesce(nopv_rooms, hotel_rooms, roomcount),
    final_rooms = case_when(
      hotel_id == "ho370651" ~ 176, # hardcoding a hotel that was incorrectly geocoded
      bbl == "1015577502" ~ 109, # hardcoding a hotel that was incorrectly geocoded
      TRUE ~ final_rooms
    )
  ) %>%
  mutate(
    bldgclass_names = case_when(
      bldgcl == "H1" ~ "Luxury Hotel",
      bldgcl == "H2" ~ "Full Service Hotel",
      bldgcl == "H3" ~ "Limited Service; Many Affiliated w/ National Chain",
      bldgcl == "H4" ~ "Motel",
      bldgcl == "H5" ~ "Hotel; Private Club, Luxury Type",
      bldgcl == "H6" ~ "Apartment Hotel",
      bldgcl == "H7" ~ "Apartment Hotel - Coop",
      bldgcl == "H8" ~ "Dormitory",
      TRUE ~ "Miscellaneous Hotel"
    ),
    main_zone = case_when(
      str_sub(zonedist1, 1, 1) == "C" ~ "Commercial",
      str_sub(zonedist1, 1, 1) == "R" ~ "Residential",
      str_sub(zonedist1, 1, 1) == "M" ~ "Manufacturing",
      zonedist1 == "BPC" ~ "Commercial"
      # Battery Park City is a special district with its own rules, but in looking at the locations
      # of the two hotels there, the subdistricts are close-enough to commercial for these purposes.
    ),
    article1_eligible = case_when(
    (main_zone == "Residential" | main_zone == "Commercial") &
      cd %in% article1_cds &
      yearbuilt < 1961 ~ TRUE,
    TRUE ~ FALSE
  )
  ) %>%
  filter(
    !bldgclass_names %in% c(
      "Apartment Hotel",
      "Apartment Hotel - Coop",
      "Hotel; Private Club, Luxury Type",
      "Dormitory"
    ),
    include != FALSE,!is.na(latitude)
  ) 

write_rds(final_hotels, "data/final_hotels.rds")

```

